---
title: "Joint paper MAGGE-pH 16S"
author: ""
date: ""
output: html_document
---
  
Data analysis for joint paper combining functional, 16S and chemical data. 

```{r global_options, echo=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.align = 'center', results = 'hide', echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
```


```{r load data}

library("phyloseq")

Magge_joint_16S <- readRDS("path/MAGGE_new_16S_physeq.rds")

sample_data(Magge_joint_16S)$Country_code <- substr(sample_data(Magge_joint_16S)$Country, 1, 2)

sample_data(Magge_joint_16S)$Average_yearly_rainfall = gsub("na", "531.93", sample_data(Magge_joint_16S)$Average_yearly_rainfall)

sample_data(Magge_joint_16S)$Average_yearly_rainfall <- as.numeric(sample_data(Magge_joint_16S)$Average_yearly_rainfall)


library(dplyr)
library(stringr)
library(microViz)


# Removing numbers from strings
replacement_string <- c("e1" = "e", "e2" = "e", "e3" = "e", "e4" = "e",
                        "l1" = "l", "l2" = "l", "l3" = "l", "l4" = "l",
                        "finely ground limestone" = "lime", "No lime" = "Control",
                        "0 t/ha lime" = "Control")

Magge_joint_16S <- Magge_joint_16S %>%
  ps_mutate(Treatment = str_replace_all(Treatment, replacement_string))

Magge_joint_16S_metadata <- data.frame(sample_data(Magge_joint_16S))



```

Rarefaction curves to  determine subsampling level. 17000 reads chosen. All samples used for analysis. Data was then subsampled to only use control sites (no pH treatment/liming)

```{r rarefaction}

# Remove any samples with less than 100 sequences
pruned <- prune_samples(sample_sums(Magge_joint_16S) >= 100, Magge_joint_16S)
pruned

library(ggplot2)
library(ranacapa)

# Calculate rarefaction curves, then plot and color by subsite
p <- ggrare(pruned, step = 1000, color = "Country", se =
              FALSE,parallel = TRUE)+
  scale_y_log10() +
  scale_x_continuous(limits = c(1, 170000))
p


p + facet_wrap(~Country)

# Rarefy at 17k

Magge_joint_16S_rare <- rarefy_even_depth(Magge_joint_16S, 17000, rngseed=TRUE)
```

NMDS ordinations based on 16s data
ANOSIM and ADONIS results shown below done by Country and then pH, with an additional Mantel test for pH at the end.

```{r Compare beta diversity pH, results='markup'}

# Ordinate data using Non-metric multidimensional scaling (NMDS) on Bray–Curtis dissimilarity (distances)
set.seed(1)
NMDS_16S.ord <- ordinate(Magge_joint_16S_rare, "NMDS", "bray")

# By calling the newly created file we can get the stress value of the plot
NMDS_16S.ord

library(vegan)
library(grid)

stressplot(NMDS_16S.ord)

# Create label for stress value
Stress_val = grobTree(textGrob("Stress = 0.16", x = 0.05,  y = 0.95, hjust = 0, gp = gpar(col = "Black", fontsize = 14,fontface = "italic")))

library(ggrepel)
library(wacolors)

options(ggrepel.max.overlaps = Inf)


Magge_16S_NMDS <- plot_ordination(Magge_joint_16S_rare, NMDS_16S.ord, type = "samples",color="Soil_pH")+
  geom_text_repel(aes(label = Country_code)) + 
  scale_color_wa_c(wacolors$vantage, reverse=TRUE)+
  geom_point(size = 1) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(color = "Soil pH")+
  annotation_custom(Stress_val)+
  theme_bw(base_size=12)



Magge_16S_NMDS


# Stats by Country
# ANOSIM
Country_group = get_variable(Magge_joint_16S_rare, "Country")
Country_ano = anosim(phyloseq::distance(Magge_joint_16S_rare, "bray"), Country_group)
Country_ano$signif
Country_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_16S_rare), "data.frame")
# Calculate your Bray distance matrix
Country_ado = phyloseq::distance(Magge_joint_16S_rare, "bray")
# Perform your ADONIS test
Country_ado_stats <- adonis2(Country_ado ~ Country, df_ado)

# Check results
Country_ado_stats


# Stats by pH
# ANOSIM
pH_group = get_variable(Magge_joint_16S_rare, "Soil_pH")
pH_ano = anosim(phyloseq::distance(Magge_joint_16S_rare, "bray"), pH_group)
pH_ano$signif
pH_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_16S_rare), "data.frame")
# Calculate your Bray distance matrix
pH_ado = phyloseq::distance(Magge_joint_16S_rare, "bray")
# Perform your ADONIS test
pH_ado_stats <- adonis2(pH_ado ~ Soil_pH, df_ado)

# Check results
pH_ado_stats

# Mantel correlation
### pH matrix
ph.dist <- dist(df_ado$Soil_pH, method = "euclidean")
dist.matrix <- as.matrix(ph.dist)

### Mantel tests
mantel(pH_ado, ph.dist, method = "spearman")

```


NMDS color based on rain

```{r NMDS color based on rain, results='markup'}

Magge_16S_NMDS_rain <- plot_ordination(Magge_joint_16S_rare, NMDS_16S.ord, type = "samples",color="Average_yearly_rainfall")+
  geom_text_repel(aes(label = Country_code)) + 
  scale_color_wa_c(wacolors$vantage, reverse=TRUE)+
  geom_point(size = 1) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(color = "Average yearly rainfall")+
  annotation_custom(Stress_val)+
  theme_bw(base_size=12)



Magge_16S_NMDS_rain


# Stats by rain
# ANOSIM
rain_group = get_variable(Magge_joint_16S_rare, "Average_yearly_rainfall")
rain_ano = anosim(phyloseq::distance(Magge_joint_16S_rare, "bray"), rain_group)
rain_ano$signif
rain_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_16S_rare), "data.frame")
# Calculate your Bray distance matrix
rain_ado = phyloseq::distance(Magge_joint_16S_rare, "bray")
# Perform your ADONIS test
rain_ado_stats <- adonis2(rain_ado ~ Average_yearly_rainfall, df_ado)

# Check results
rain_ado_stats

# Mantel correlation
### Rain matrix
rain.dist <- dist(df_ado$Average_yearly_rainfall, method = "euclidean")
dist.matrix <- as.matrix(rain.dist)

### Mantel tests
mantel(rain_ado, rain.dist, method = "spearman")
```

PCA analysis based on chemical data

```{r chem PCA, results='markup'}
# Create PCA including all soil properties

library("tidyverse")

# Reorganize data and convert to numerical

Magge_joint_16S_metadata <- Magge_joint_16S_metadata %>% relocate(c(Treatment, Soil_type, Soil_depth, Cumulative_N20, Cumulative_N2, N2O_ratio, nirk.s:nos.nir), .before = Note)

# Replace NA with missing value 
Magge_joint_16S_metadata["Average_yearly_rainfall"][is.na(Magge_joint_16S_metadata["Average_yearly_rainfall"])] <- 531.93

Magge_joint_16S_metadata <- Magge_joint_16S_metadata %>% mutate_at(c(23:51), as.numeric)


# PCA plot

chem.pca <- prcomp(Magge_joint_16S_metadata[,c(23:42)], scale = TRUE)
summary(chem.pca)
var_explained <- chem.pca$sdev^2/sum(chem.pca$sdev^2)
var_explained[1:5]

library(ggfortify)
autoplot(chem.pca, data = Magge_joint_16S_metadata, color="Soil_pH",loadings = TRUE, loadings.colour = 'red',
         loadings.label = TRUE, loadings.label.size = 3,loadings.label.repel=T,box.padding = 1)+ 
  geom_text_repel(aes(label = Country_code, color = Soil_pH)) + 
  geom_point(size=1, aes(color = Soil_pH))+
  scale_color_wa_c(wacolors$vantage, reverse=TRUE)+
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size = 12), 
        axis.text.y = element_text(colour = "black", size = 12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  labs(color = "Soil pH") + theme_bw(base_size = 12)


# Add table for factor loading regressions
chem.pca.correlation <- data.frame(chem.pca$rotation[,1:2])
chem.pca.correlation

# Replot colored by rain
autoplot(chem.pca, data = Magge_joint_16S_metadata, color="Average_yearly_rainfall",loadings = TRUE, loadings.colour = 'red',
         loadings.label = TRUE, loadings.label.size = 3,loadings.label.repel=T,box.padding = 1)+ 
  geom_text_repel(aes(label = Country_code, color = Average_yearly_rainfall)) + 
  geom_point(size=1, aes(color = Average_yearly_rainfall)) +
  scale_color_wa_c(wacolors$vantage, reverse = TRUE) +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size = 12), 
        axis.text.y = element_text(colour = "black", size = 12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  labs(color = "Average yearly rainfall") + theme_bw(base_size = 12)


```


PCA analysis based on qPCR data
Notice the New Zealand outlier skewing the plot. Seems very strange so should we remove?

```{r Func PCA, results='markup'}


# PCA plot

Func.pca <- prcomp(Magge_joint_16S_metadata[,c(43:51)], scale = TRUE)
summary(Func.pca)
var_explained <- Func.pca$sdev^2/sum(Func.pca$sdev^2)
var_explained[1:5]

autoplot(Func.pca, data = Magge_joint_16S_metadata, color="Soil_pH",loadings = TRUE, loadings.colour = 'red',
         loadings.label = TRUE, loadings.label.size = 3,loadings.label.repel = T,box.padding = 1)+ 
  geom_text_repel(aes(label = Country_code, color = Soil_pH)) + 
  geom_point(size = 1, aes(color = Soil_pH))+
  scale_color_wa_c(wacolors$vantage, reverse = TRUE) +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  labs(color = "Soil pH") + theme_bw(base_size=12)


# Add table for factor loading regressions
Func.pca.correlation<-data.frame(Func.pca$rotation[,1:2])
Func.pca.correlation

# Replot and color by rain
autoplot(Func.pca, data = Magge_joint_16S_metadata, color="Average_yearly_rainfall",loadings = TRUE, loadings.colour = 'red',
         loadings.label = TRUE, loadings.label.size = 3,loadings.label.repel=T,box.padding = 1) + 
  geom_text_repel(aes(label = Country_code, color = Average_yearly_rainfall)) + 
  geom_point(size = 1, aes(color = Average_yearly_rainfall))+
  scale_color_wa_c(wacolors$vantage, reverse = TRUE)+
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(color = "Average yearly rainfall") + theme_bw(base_size=12)


```

Stacked bar plot based on phylum (draft and clean version)

```{r stack bar phylum}

library(plyr)
library(scales)

# Clean graph and cluster by country

library(dplyr)
library(microViz)
library(fantaxtic)
library(speedyseq)
library(phylosmith)

Magge_joint_16S_rare <- merge_treatments(Magge_joint_16S_rare, c("Country", "Treatment"))

MJ16SR_clean <- name_na_taxa(Magge_joint_16S_rare, na_label = "Unidentified <tax> (<rank>)")

MJ16SR_clean2 <- MJ16SR_clean %>%
  ps_select(Country, Treatment, Country_Treatment, Soil_pH) %>% # avoids lots of phyloseq::merge_samples warnings
  merge_samples2(group = "Country_Treatment", funs = list()) %>%
  tax_glom(taxrank = "Phylum")

MJ16SR_clean2 %>%
  tax_filter(
    tax_level = "Phylum", min_prevalence = 0.001,
    prev_detection_threshold = 100
  ) %>%
  ps_mutate(
    Treatment = factor(Treatment, levels = c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)','Dolomite', 'Larvikite', 'Marble','Norite','Olivine','10 ton/ha lime', '20 ton /ha lime','Slaked lime', 'Tunnel kiln slag', 'Mixed lime'))) %>%  
  ps_mutate(
    Country = factor(Country, levels = c('Denmark', 'Ireland', 'Finland', 'Norway', 'New_Zealand', 'France_LQ', 'Sweden_B', 'France_TX', 'Sweden_LH'))) %>%  
  comp_barplot(
    tax_level = "Phylum", n_taxa = 10,
    bar_outline_colour = NA,
    bar_width = 0.7,
    taxon_renamer = toupper,
    facet_by = "Country",
    label = "Treatment",
    x = "Treatment"
    ) + 
  theme(legend.text = element_text(size = 12),
        legend.title =element_text(size = 14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 10, vjust = 1, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())


```

Stacked bar plot based on Order (clean version)

```{r stack bar Order}

# Clean graph and cluster by country

MJ16SR_clean2_Order<-MJ16SR_clean%>%
  ps_select(Country, Treatment, Country_Treatment, Soil_pH) %>% # avoids lots of phyloseq::merge_samples warnings
  merge_samples2(group = "Country_Treatment", funs = list())%>%
  tax_glom(taxrank = "Order")

MJ16SR_clean2_Order %>%
  tax_filter(
    tax_level = "Order", min_prevalence = 0.001,
    prev_detection_threshold = 100
  ) %>%
  ps_mutate(
    Treatment = factor(Treatment, levels = c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)','Dolomite', 'Larvikite', 'Marble','Norite','Olivine','10 ton/ha lime', '20 ton /ha lime','Slaked lime', 'Tunnel kiln slag', 'Mixed lime'))) %>%  
  ps_mutate(
    Country = factor(Country, levels = c('Denmark', 'Ireland', 'Finland', 'Norway', 'New_Zealand', 'France_LQ', 'Sweden_B', 'France_TX', 'Sweden_LH'))) %>%  
  comp_barplot(
    tax_level = "Order", n_taxa = 15,
    bar_outline_colour = NA,
    bar_width = 0.7,
    taxon_renamer = toupper,
    facet_by = "Country",
    label = "Treatment",
    x = "Treatment"
  ) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(angle=45, colour = "black", size=10, vjust = 1, hjust = 1), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())


```


Lime treatment comparison for functional genes

```{r lime treatment comparison functional genes}

# Read the file again because it has been reorganized
Magge_joint_16S_metadata <- data.frame(sample_data(Magge_joint_16S))

func_data_long <- gather(Magge_joint_16S_metadata, Target, measurement, bac:coma, factor_key = TRUE)
func_data_long


func_data_long <- func_data_long[ -c(1:5, 7, 11:42) ]
func_data_long
func_data_long <- func_data_long %>% mutate_at(c('measurement'), as.numeric)

library(extrafont)
library(grDevices)
windowsFonts(Time = windowsFont("Times"))

library(ggpubr)

Denmark_order <- c('Control', '4 t/ha lime', '8 t/ha lime', '12 t/ha lime') 
func_data_long_De <- func_data_long

func_data_long_De$Treatment = factor(func_data_long_De$Treatment, levels = c('Control', '4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))

DE_comparisons <- list(c('Control', '4 t/ha lime'), c('Control', '8 t/ha lime'), c('Control', '12 t/ha lime'))


# Denmark results
ggplot(subset(func_data_long_De, Country %in% c("Denmark")), aes(x = Treatment, y = measurement))+ 
  stat_compare_means(method = "kruskal.test", size = 3, aes(group = Treatment), label.x.npc = "left", vjust = -9) +
  geom_boxplot() +
  facet_wrap(.~Target, drop = TRUE, scales = "free_y")+
  theme(text = element_text(family = "Times"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 8, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 8),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  ggtitle("Denmark") +
  stat_compare_means(method = "t.test" , comparisons = DE_comparisons, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.4)))


Ireland_order <- c('Control', 'Lime (2011, 2013)', 'Lime (2013)') 
func_data_long_IR <- func_data_long

func_data_long_IR$Treatment = factor(func_data_long_IR$Treatment, levels = c('Control', 'Lime (2011, 2013)', 'Lime (2013)'))
func_data_long_IR <- func_data_long_IR %>% 
  drop_na(c('Treatment'))

IR_comparisons <- list(c('Control', 'Lime (2011, 2013)'), c('Control', 'Lime (2013)'))

# Ireland results
ggplot(func_data_long_IR, aes(x = Treatment, y = measurement))+ 
  stat_compare_means(method = "kruskal.test", size = 3, aes(group = Treatment), label.x.npc = "left", vjust = -9) +
  geom_boxplot()+
  facet_wrap(.~Target, drop = TRUE, scales = "free_y")+
  theme(text = element_text(family = "Times"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 8, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 8),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  ggtitle("Ireland")+
  stat_compare_means(method = "t.test" , comparisons = IR_comparisons, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.4)))


# Norway results
NO_comparisons <- list(c('Control', 'Dolomite'), c('Control', 'Larvikite'), c('Control', 'Marble'), c('Control', 'Norite'), c('Control', 'Olivine'))


ggplot(subset(func_data_long, Country %in% c("Norway")), aes(x=Treatment, y=measurement)) + 
  geom_boxplot()+
  facet_wrap(.~Target, drop = TRUE, scales = "free_y")+
  theme(text = element_text(family = "Times"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 8, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 8),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  stat_compare_means(method = "kruskal.test", size = 3, label.x.npc = "left", vjust = -9)+
  ggtitle("Norway")+
  stat_compare_means(method = "t.test" , size = 3, comparisons = NO_comparisons) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.4)))



SWLH_comparisons <- list(c('Control', 'Mixed lime'), c('Control', 'Slaked lime'), c('Control', 'Tunnel kiln slag'))

# Sweden_LH results
ggplot(subset(func_data_long, Country %in% c("Sweden_LH")), aes(x=Treatment, y=measurement)) + 
  geom_boxplot()+
  facet_wrap(.~Target, drop = TRUE, scales = "free_y")+
  theme(text = element_text(family = "Times"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face="bold"),
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 8, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 8),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  stat_compare_means(method = "kruskal.test", size = 3, label.x.npc = "left", vjust = -9)+
  ggtitle("Sweden_LH")+
  stat_compare_means(method = "t.test", size = 3, comparisons = SWLH_comparisons) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.4)))


SWB_comparisons <- list( c('Control', '10 ton/ha lime'), c('Control', '20 ton /ha lime'))


SwedenB_order <- c('Control', '10 ton/ha lime', '20 ton /ha lime') 
func_data_long_SwB <- func_data_long

func_data_long_SwB$Treatment = factor(func_data_long_SwB$Treatment, levels=c('Control', '10 ton/ha lime', '20 ton /ha lime'))
func_data_long_SwB<-func_data_long_SwB %>% 
  drop_na(c('Treatment'))

# Sweden_B results
ggplot(func_data_long_SwB, aes(x=Treatment, y=measurement)) + 
  geom_boxplot()+
  facet_wrap(.~Target, drop = TRUE, scales = "free_y")+
  theme(text = element_text(family = "Times"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face="bold"),
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 8, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 8),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  stat_compare_means(method = "kruskal.test", size = 3, label.x.npc = "left", vjust = -9)+
  ggtitle("Sweden_B")+
  stat_compare_means(method = "t.test", size = 3, comparisons = SWB_comparisons) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.4)))



```


distance comparisons Ireland

```{r distance comparisons Ireland}

Magge_joint_16S_rare_IR <- subset_samples(Magge_joint_16S_rare, Country == "Ireland")
Magge_joint_16S_rare_IR

Totu_table = t(otu_table(Magge_joint_16S_rare_IR)) # transpose otu table
otu_table(Magge_joint_16S_rare_IR) = Totu_table

p = Magge_joint_16S_rare_IR
m = "bray"
s = "X.SampleID"
d = "Treatment"   # Day

# Calculate distances
library(reshape2)
library(reshape)
library(dplyr)
wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor, as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor, as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_16s_ireland = wu.sd %>%
  filter(Treatment == 'Control') %>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Ireland")


library(ggpubr)

# Order factors
wu.sd.sums_16s_ireland$Treatment2 = factor(wu.sd.sums_16s_ireland$Treatment2, levels = c('Control','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)'))

IR_16S_dist_plot<-ggplot(wu.sd.sums_16s_ireland, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size = 18), axis.title.y = element_text(size = 18), 
        strip.text.x = element_text(face = "bold", size = 18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size = 3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control") + ylim(0, 1)+ggtitle("Ireland")

IR_16S_dist_plot

```

distance comparisons Norway
```{r distance comparisons Norway}

Magge_joint_16S_rare_NO <- subset_samples(Magge_joint_16S_rare, Country == "Norway")
Magge_joint_16S_rare_NO

Totu_table =t(otu_table(Magge_joint_16S_rare_NO)) #transpose otu table
otu_table(Magge_joint_16S_rare_NO)=Totu_table

p = Magge_joint_16S_rare_NO
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calc distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_16s_norway = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Norway")


# Order factors
wu.sd.sums_16s_norway$Treatment2 = factor(wu.sd.sums_16s_norway$Treatment2, levels=c('Control','Dolomite', 'Larvikite', 'Marble','Norite','Olivine'))

NO_16S_dist_plot<-ggplot(wu.sd.sums_16s_norway, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Norway")
NO_16S_dist_plot


```

distance comparisons Denmark

```{r distance comparisons Denmark}

Magge_joint_16S_rare_DE <- subset_samples(Magge_joint_16S_rare, Country == "Denmark")
Magge_joint_16S_rare_DE

Totu_table =t(otu_table(Magge_joint_16S_rare_DE)) #transpose otu table
otu_table(Magge_joint_16S_rare_DE)=Totu_table

p = Magge_joint_16S_rare_DE
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculating distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Removing self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_16s_denmark = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Denmark")


# Order factors
wu.sd.sums_16s_denmark$Treatment2 = factor(wu.sd.sums_16s_denmark$Treatment2, levels=c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))

DE_16S_dist_plot<-ggplot(wu.sd.sums_16s_denmark, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Denmark")

DE_16S_dist_plot

```


distance comparisons Sweden_LH

```{r distance comparisons Sweden_LH}

Magge_joint_16S_rare_SWLH <- subset_samples(Magge_joint_16S_rare, Country == "Sweden_LH")
Magge_joint_16S_rare_SWLH

Totu_table =t(otu_table(Magge_joint_16S_rare_SWLH)) #transpose otu table
otu_table(Magge_joint_16S_rare_SWLH)=Totu_table

p = Magge_joint_16S_rare_SWLH
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculate distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_16s_swedenlh = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Sweden_LH")


# Order factors
# wu.sd.sums$Treatment2 = factor(wu.sd.sums$Treatment2, levels=c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))

SWLH_16S_dist_plot<-ggplot(wu.sd.sums_16s_swedenlh, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30,  aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Sweden_LH")

SWLH_16S_dist_plot



```


distance comparisons Sweden_B

```{r distance comparisons Sweden_B}

Magge_joint_16S_rare_SWB <- subset_samples(Magge_joint_16S_rare, Country == "Sweden_B")
Magge_joint_16S_rare_SWB

Totu_table =t(otu_table(Magge_joint_16S_rare_SWB)) #transpose otu table
otu_table(Magge_joint_16S_rare_SWB)=Totu_table

p = Magge_joint_16S_rare_SWB
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculate distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_16s_swedenb = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Sweden_B")


# Order factors
wu.sd.sums_16s_swedenb$Treatment2 = factor(wu.sd.sums_16s_swedenb$Treatment2, levels=c('Control','10 ton/ha lime', '20 ton /ha lime'))

SWB_16S_dist_plot<-ggplot(wu.sd.sums_16s_swedenb, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30,  aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Sweden_B")

SWB_16S_dist_plot


#combine all plots
library(gridExtra)
#pdf("/Volumes/micro-shared$/MoralesLab/Manuscripts/Magge-pH combined microbiome/Finals/figs/16S_distances_pH.pdf", width = 12, height = 16)

# Open a new pdf file
#grid.arrange(NO_16S_dist_plot,DE_16S_dist_plot,SWB_16S_dist_plot,SWLH_16S_dist_plot,IR_16S_dist_plot, 
#            ncol=2, nrow=3, widths=c(4, 4), heights=c(2,2,2))
#dev.off()
    
```


```{r making a summary table, results='markup'}

library(dplyr)
library(broom)
library(tidyr)
library(rstatix)

# Filter countries
Magge_joint_16S_metadata_filtered <- Magge_joint_16S_metadata %>%
  select(X.SampleID, Country, Soil_pH, Treatment, 16:50) %>%
  filter(Country %in% c("Denmark", "Sweden_LH", "Sweden_B", "Norway", "Ireland"))

# Gathering
Magge_joint_16S_metadata_filtered_gathered <- Magge_joint_16S_metadata_filtered %>%
  relocate(Treatment, .before = Soil_pH) %>%
  gather(Parameter, Measurement, Soil_pH:nosi.nosii, factor_key = TRUE)
  
# Performing kruskal test comparing all parameters to treatment
Magge_joint_16S_metadata_filtered_kw <- Magge_joint_16S_metadata_filtered_gathered %>%
  group_by(Parameter, Country) %>%
  do(tidy(kruskal.test(.$Measurement, .$Treatment))) %>%
  ungroup() %>%
  arrange(Country)
  
# Calculating min and max values
Magge_joint_16S_metadata_filtered_extreme <- Magge_joint_16S_metadata_filtered_gathered %>%
  group_by(Parameter, Country) %>%
  summarise_at(vars(Measurement), funs(min, max)) %>%
  arrange(Country)

# Calculating effect size
Magge_joint_16S_metadata_filtered_effect_size <- Magge_joint_16S_metadata_filtered_gathered %>%
  group_by(Parameter, Country) %>%
  kruskal_effsize(Measurement ~ Treatment) %>%
  arrange(Country)
 
# Joining tables
Magge_joint_16S_metadata_filtered_joined <- bind_cols(Magge_joint_16S_metadata_filtered_kw, Magge_joint_16S_metadata_filtered_extreme, Magge_joint_16S_metadata_filtered_effect_size)
Magge_joint_16S_metadata_filtered_joined_named <- Magge_joint_16S_metadata_filtered_joined %>%
  select(Parameter...1, Country...2, p.value, effsize, min, max) %>%
  dplyr::rename("Parameter" = Parameter...1, "Country" = Country...2, "P_value" = p.value, "Effect_size" = effsize, "Minimal value" = min, "Maximal value" = max)

# Filtering for significant values only
Magge_joint_16S_metadata_filtered_joined_named_significant <- Magge_joint_16S_metadata_filtered_joined_named %>%
  filter(P_value < 0.05)
Magge_joint_16S_metadata_filtered_joined_named_significant

# writexl::write_xlsx(Magge_joint_16S_metadata_filtered_joined_named_significant, "MAGGE_Summary_table.xlsx")
```



```{r Find significantly affected Phyla using rstatix}

#no hits after p adjusted so used just p value
Magge_joint_16S_rare_DE_df<-Magge_joint_16S_rare_DE %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 
#how does this community look?
d<- ggplot(Magge_joint_16S_rare_DE_df,aes(x=Treatment,y=Abundance,fill=Phylum))+ 
  geom_bar(stat="identity",position="fill") +
  scale_y_continuous()+
  ylab("Relative Abundance")
d

library(rstatix)
# do tests
KW_taxa_DE<-Magge_joint_16S_rare_DE_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Denmark")



# make a list of the significant genera
only_sig_taxa_DE<-Magge_joint_16S_rare_DE_df %>% 
  filter(OTU %in% KW_taxa_DE$OTU) 


only_sig_taxa_DE$Treatment = factor(only_sig_taxa_DE$Treatment, levels=c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))

# plot
DE_sig_tax_plot<-ggplot(only_sig_taxa_DE, aes(x=Treatment, y=Abundance, fill=Phylum)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ggtitle("Denmark")

DE_sig_tax_plot



#Ireland
#no hits after p adjusted so used just p value
Magge_joint_16S_rare_IR_df<-Magge_joint_16S_rare_IR %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_IR<-Magge_joint_16S_rare_IR_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Ireland")



# make a list of the significant genera
only_sig_taxa_IR<-Magge_joint_16S_rare_IR_df %>% 
  filter(OTU %in% KW_taxa_IR$OTU) 


only_sig_taxa_IR$Treatment = factor(only_sig_taxa_IR$Treatment, levels = c('Control','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)'))

# plot
IR_sig_tax_plot<-ggplot(only_sig_taxa_IR, aes(x=Treatment, y=Abundance, fill=Phylum)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ggtitle("Ireland")

IR_sig_tax_plot


#Sweden LH
#no hits after p adjusted so used just p value
Magge_joint_16S_rare_SWLH_df<-Magge_joint_16S_rare_SWLH %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_SWLH<-Magge_joint_16S_rare_SWLH_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Sweden_LH")



# make a list of the significant genera
only_sig_taxa_SWLH<-Magge_joint_16S_rare_SWLH_df %>% 
  filter(OTU %in% KW_taxa_SWLH$OTU) 


only_sig_taxa_SWLH$Treatment = factor(only_sig_taxa_SWLH$Treatment, levels=c('Control','Mixed lime', 'Slaked lime', 'Tunnel kiln slag'))

# plot
SWLH_sig_tax_plot<-ggplot(only_sig_taxa_SWLH, aes(x=Treatment, y=Abundance, fill=Phylum, group=Treatment)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = 'Control')+ggtitle("Sweden LH")

SWLH_sig_tax_plot




#no hits after p adjusted so used just p value
Magge_joint_16S_rare_SWB_df<-Magge_joint_16S_rare_SWB %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_SWB<-Magge_joint_16S_rare_SWB_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Sweden_B")



# make a list of the significant genera
only_sig_taxa_SWB<-Magge_joint_16S_rare_SWB_df %>% 
  filter(OTU %in% KW_taxa_SWB$OTU) 


only_sig_taxa_SWB$Treatment = factor(only_sig_taxa_SWB$Treatment, levels=c('Control','10 ton/ha lime', '20 ton /ha lime'))

# plot
SWB_sig_tax_plot<-ggplot(only_sig_taxa_SWB, aes(x=Treatment, y=Abundance, fill=Phylum, group=Treatment)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = 'Control')+ggtitle("Sweden LH")

SWB_sig_tax_plot



# Combined table for taxa
KW_taxa_all <- bind_rows(KW_taxa_DE, KW_taxa_IR, KW_taxa_SWB, KW_taxa_SWLH)
KW_taxa_all <- KW_taxa_all %>%
  select(-.y., -n, -df) %>%
  arrange(desc(p))

KW_taxa_all

```

```{r combined distance plots}

# For that run both Joint_MAGGE_16S_all_rev.Rmd and Joint_MAGGE_ITS_all_rev.Rmd

wu.sd.sums_16s_combined <- bind_rows(wu.sd.sums_16s_denmark, wu.sd.sums_16s_ireland, wu.sd.sums_16s_norway, wu.sd.sums_16s_swedenb, wu.sd.sums_16s_swedenlh)
wu.sd.sums_16s_combined$type <- rep("16S", 269)
wu.sd.sums_its_combined <- bind_rows(wu.sd.sums_its_denmark, wu.sd.sums_its_ireland, wu.sd.sums_its_norway, wu.sd.sums_its_swedenb, wu.sd.sums_its_swedenlh)
wu.sd.sums_its_combined$type <- rep("ITS", 269)
wu.sd.sums_all_combined <- bind_rows(wu.sd.sums_16s_combined, wu.sd.sums_its_combined)


wu.sd.sums_all_combined$Treatment2 = factor(wu.sd.sums_all_combined$Treatment2, levels = c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime',
                                                                                            'Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)',
                                                                                              'Dolomite', 'Larvikite','Marble','Norite','Olivine',
                                                                                              '10 ton/ha lime', '20 ton /ha lime', 'Mixed lime',
                                                                                              'Slaked lime', 'Tunnel kiln slag'))

stat.test <- wu.sd.sums_all_combined %>%
  group_by(country) %>%
  t_test(value ~ Treatment2, ref.group = "Control") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance()
stat.test 

stat.test <- stat.test %>%
 add_xy_position(x = "Treatment2")


kw <- wu.sd.sums_all_combined %>%
  group_by(type, country) %>%
  do(tidy(kruskal.test(.$value, .$Treatment2))) %>%
  mutate(across(where(is.numeric), round, 6))

kw16 <- filter(kw, type == "16S")
kwits <- filter(kw, type == "ITS")


distance_plot <- ggplot(wu.sd.sums_all_combined, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_boxplot(aes(fill = type)) + labs(y = "Bray distance", x = "", fill = "Sequence type") +
  facet_wrap(~country, drop = TRUE, scales = "free") + scale_fill_viridis_d() +
  theme(
        axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size = 8, angle = 45),
        axis.text.y = element_text(colour = "black", size = 8),
        axis.title.y = element_text(size = 10), 
        strip.text.x = element_text(face = "bold", size = 12),
        legend.position = c(0.8, 0.2),
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  scale_x_discrete(breaks = wu.sd.sums_all_combined$Treatment2) +
  #stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.01, hide.ns = TRUE)
  #stat_compare_means(method = "kruskal.test", size = 2, vjust = -1)
  geom_text(data = kw16, size = 3, aes(x = 1.3, y = 1.1, label = paste0("Kruskal-Wallis\n p (16S) =", p.value))) +
  geom_text(data = kwits, size = 3, aes(x = 1.3, y = 1, label = paste0("p (ITS) =", p.value)))

distance_plot


```

```{r response ratio calculation and LRR plot for Denmark, echo=FALSE,results='hide',fig.keep='all'}

library(SingleCaseES)

# Writing functions for LRR
LRR_MAGGE <- function(x = Magge_joint_16S_metadata_filtered_gathered, country, treatment, parameter) {

Controls <- Magge_joint_16S_metadata_filtered_gathered %>%
  filter(Country == country) %>%
  select(Measurement, Treatment, Parameter) %>%
  filter(Treatment == "Control" & Parameter == parameter) %>%
  select(Measurement)

Response <- Magge_joint_16S_metadata_filtered_gathered %>%
  filter(Country == country) %>%
  select(Measurement, Treatment, Parameter) %>%
  filter(Treatment == treatment & Parameter == parameter) %>%
  select(Measurement)

LRRd(as.numeric(Controls$Measurement), as.numeric(Response$Measurement))

}

Denmark_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Denmark", "4 t/ha lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Denmark", "8 t/ha lime", y),
            LRR_MAGGE(x, "Denmark", "12 t/ha lime", y))
}

# Making a table with results of LRR
Denmark_LRR_result <- bind_rows(
          Denmark_LRR(y ="Soil_pH"),
          Denmark_LRR(y = "soc"),
          Denmark_LRR(y = "tn"),
          Denmark_LRR(y = "sand"),
          Denmark_LRR(y = "silt"),
          Denmark_LRR(y = "clay"),
          Denmark_LRR(y = "Al"),
          Denmark_LRR(y = "Ca"),
          Denmark_LRR(y = "Co"),
          Denmark_LRR(y = "Cu"),
          Denmark_LRR(y = "Fe"),
          Denmark_LRR(y = "K"),
          Denmark_LRR(y = "Mg"),
          Denmark_LRR(y = "Mn"),
          Denmark_LRR(y = "Na"),
          Denmark_LRR(y = "P"),
          Denmark_LRR(y = "S"),
          Denmark_LRR(y = "Zn"),
          Denmark_LRR(y = "its")
          )

# Adding columns
Denmark_LRR_result$Parameter <- as.factor(rep(c("Soil_pH", "soc", "tn", "sand", "silt", "clay", "Al", "Ca", "Co", "Cu", "Fe",
                                       "K", "Mg", "Mn", "Na", "P", "S", "Zn", "its"), each = 3))

Denmark_LRR_result$Treatment <- as.factor(rep(c("4 t/ha lime", "8 t/ha lime", "12 t/ha lime")))
Denmark_LRR_result$Treatment <- factor(Denmark_LRR_result$Treatment, levels = c("4 t/ha lime", "8 t/ha lime", "12 t/ha lime"))
Denmark_LRR_result$Country <- rep(c("Denmark"))

# Removing not significant values
Denmark_LRR_result <- Denmark_LRR_result %>%
  filter(Est != 0)

library(ggplot2)
library(wacolors)
library(extrafont)


# LRR plot                      
Denmark_LRR_result_plot <- ggplot(Denmark_LRR_result, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Denmark")
  

#Denmark_LRR_result_plot


```

```{r response ratio calculation and LRR plot for Ireland, echo=FALSE,results='hide',fig.keep='all'}

Ireland_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Ireland", "Lime (2011)", y), # Change country and treatment here
            LRR_MAGGE(x, "Ireland", "Lime (2011, 2013)", y),
            LRR_MAGGE(x, "Ireland", "Lime (2013)", y))
}

# Making a table with results of LRR
Ireland_LRR_result <- bind_rows(
          Ireland_LRR(y ="Soil_pH"),
          Ireland_LRR(y = "soc"),
          Ireland_LRR(y = "tn"),
          Ireland_LRR(y = "sand"),
          Ireland_LRR(y = "silt"),
          Ireland_LRR(y = "clay"),
          Ireland_LRR(y = "Al"),
          Ireland_LRR(y = "Ca"),
          Ireland_LRR(y = "Co"),
          Ireland_LRR(y = "Cu"),
          Ireland_LRR(y = "Fe"),
          Ireland_LRR(y = "K"),
          Ireland_LRR(y = "Mg"),
          Ireland_LRR(y = "Mn"),
          Ireland_LRR(y = "Na"),
          Ireland_LRR(y = "P"),
          Ireland_LRR(y = "S"),
          Ireland_LRR(y = "Zn"),
          Ireland_LRR(y = "its")
          )

# Adding columns
Ireland_LRR_result$Parameter <- as.factor(rep(c("Soil_pH", "soc", "tn", "sand", "silt", "clay", "Al", "Ca", "Co", "Cu", "Fe",
                                       "K", "Mg", "Mn", "Na", "P", "S", "Zn", "its"), each = 3))


Ireland_LRR_result$Treatment <- as.factor(rep(c("Lime (2011)", "Lime (2011, 2013)", "Lime (2013)")))
Ireland_LRR_result$Treatment <- factor(Ireland_LRR_result$Treatment, levels = c("Lime (2011)", "Lime (2011, 2013)", "Lime (2013)"))
Ireland_LRR_result$Country <- rep(c("Ireland"))
Ireland_LRR_result <- Ireland_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Ireland_LRR_result_plot <- ggplot(Ireland_LRR_result, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Ireland")
  

Ireland_LRR_result_plot
```

```{r response ratio calculation and LRR plot for Norway, echo=FALSE,results='hide',fig.keep='all'}
Norway_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Norway", "Dolomite", y), # Change country and treatment here
            LRR_MAGGE(x, "Norway", "Larvikite", y),
            LRR_MAGGE(x, "Norway", "Marble", y),
            LRR_MAGGE(x, "Norway", "Norite", y),
            LRR_MAGGE(x, "Norway", "Olivine", y))
}

# Making a table with results of LRR
Norway_LRR_result <- bind_rows(
          Norway_LRR(y ="Soil_pH"),
          Norway_LRR(y = "soc"),
          Norway_LRR(y = "tn"),
          Norway_LRR(y = "sand"),
          Norway_LRR(y = "silt"),
          Norway_LRR(y = "clay"),
          Norway_LRR(y = "Al"),
          Norway_LRR(y = "Ca"),
          Norway_LRR(y = "Co"),
          Norway_LRR(y = "Cu"),
          Norway_LRR(y = "Fe"),
          Norway_LRR(y = "K"),
          Norway_LRR(y = "Mg"),
          Norway_LRR(y = "Mn"),
          Norway_LRR(y = "Na"),
          Norway_LRR(y = "P"),
          Norway_LRR(y = "S"),
          Norway_LRR(y = "Zn"),
          Norway_LRR(y = "its")
          )

# Adding columns
Norway_LRR_result$Parameter <- as.factor(rep(c("Soil_pH", "soc", "tn", "sand", "silt", "clay", "Al", "Ca", "Co", "Cu", "Fe",
                                       "K", "Mg", "Mn", "Na", "P", "S", "Zn", "its"), each = 5))


Norway_LRR_result$Treatment <- as.factor(rep(c("Dolomite", "Larvikite", "Marble", "Norite", "Olivine")))
Norway_LRR_result$Treatment <- factor(Norway_LRR_result$Treatment, levels = c("Dolomite", "Larvikite", "Marble", "Norite", "Olivine"))
Norway_LRR_result$Country <- rep(c("Norway"))
Norway_LRR_result <- Norway_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Norway_LRR_result_plot <- ggplot(Norway_LRR_result, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Norway")
  

Norway_LRR_result_plot
```

```{r response ratio calculation and LRR plot for Sweden_B, echo=FALSE,results='hide',fig.keep='all'}

Sweden_B_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Sweden_B", "10 ton/ha lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Sweden_B", "20 ton /ha lime", y),
            )
}

# Making a table with results of LRR
Sweden_B_LRR_result <- bind_rows(
          Sweden_B_LRR(y ="Soil_pH"),
          Sweden_B_LRR(y = "soc"),
          Sweden_B_LRR(y = "tn"),
          Sweden_B_LRR(y = "sand"),
          Sweden_B_LRR(y = "silt"),
          Sweden_B_LRR(y = "clay"),
          Sweden_B_LRR(y = "Al"),
          Sweden_B_LRR(y = "Ca"),
          Sweden_B_LRR(y = "Co"),
          Sweden_B_LRR(y = "Cu"),
          Sweden_B_LRR(y = "Fe"),
          Sweden_B_LRR(y = "K"),
          Sweden_B_LRR(y = "Mg"),
          Sweden_B_LRR(y = "Mn"),
          Sweden_B_LRR(y = "Na"),
          Sweden_B_LRR(y = "P"),
          Sweden_B_LRR(y = "S"),
          Sweden_B_LRR(y = "Zn"),
          Sweden_B_LRR(y = "its")
          )

# Adding columns
Sweden_B_LRR_result$Parameter <- as.factor(rep(c("Soil_pH", "soc", "tn", "sand", "silt", "clay", "Al", "Ca", "Co", "Cu", "Fe",
                                       "K", "Mg", "Mn", "Na", "P", "S", "Zn", "its"), each = 2))


Sweden_B_LRR_result$Treatment <- as.factor(rep(c("10 ton/ha lime", "20 ton/ha lime")))
Sweden_B_LRR_result$Treatment <- factor(Sweden_B_LRR_result$Treatment, levels = c("10 ton/ha lime", "20 ton/ha lime"))
Sweden_B_LRR_result$Country <- rep(c("Sweden_B"))
Sweden_B_LRR_result <- Sweden_B_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Sweden_B_LRR_result_plot <- ggplot(Sweden_B_LRR_result, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white")) +
  ggtitle("Sweden B")
  

Sweden_B_LRR_result_plot
```

```{r response ratio calculation and LRR plot for Sweden_LH, echo=FALSE,results='hide',fig.keep='all'}

Sweden_LH_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Sweden_LH", "Mixed lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Sweden_LH", "Slaked lime", y),
            LRR_MAGGE(x, "Sweden_LH", "Tunnel kiln slag", y)
            )
}

# Making a table with results of LRR
Sweden_LH_LRR_result <- bind_rows(
          Sweden_LH_LRR(y ="Soil_pH"),
          Sweden_LH_LRR(y = "soc"),
          Sweden_LH_LRR(y = "tn"),
          Sweden_LH_LRR(y = "sand"),
          Sweden_LH_LRR(y = "silt"),
          Sweden_LH_LRR(y = "clay"),
          Sweden_LH_LRR(y = "Al"),
          Sweden_LH_LRR(y = "Ca"),
          Sweden_LH_LRR(y = "Co"),
          Sweden_LH_LRR(y = "Cu"),
          Sweden_LH_LRR(y = "Fe"),
          Sweden_LH_LRR(y = "K"),
          Sweden_LH_LRR(y = "Mg"),
          Sweden_LH_LRR(y = "Mn"),
          Sweden_LH_LRR(y = "Na"),
          Sweden_LH_LRR(y = "P"),
          Sweden_LH_LRR(y = "S"),
          Sweden_LH_LRR(y = "Zn"),
          Sweden_LH_LRR(y = "its")
          )

# Adding columns

Sweden_LH_LRR_result$Parameter <- as.factor(rep(c("Soil_pH", "soc", "tn", "sand", "silt", "clay", "Al", "Ca", "Co", "Cu", "Fe",
                                       "K", "Mg", "Mn", "Na", "P", "S", "Zn", "its"), each = 3))


Sweden_LH_LRR_result$Treatment <- as.factor(rep(c("Mixed lime", "Slaked lime", "Tunnel kiln slag")))
Sweden_LH_LRR_result$Treatment <- factor(Sweden_LH_LRR_result$Treatment, levels = c("Mixed lime", "Slaked lime", "Tunnel kiln slag"))
Sweden_LH_LRR_result$Country <- rep(c("Sweden_LH"))
Sweden_LH_LRR_result <- Sweden_LH_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Sweden_LH_LRR_result_plot <- ggplot(Sweden_LH_LRR_result, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Sweden LH")
  

Sweden_LH_LRR_result_plot
```


```{r treatments as pH, echo=FALSE,results='hide',fig.keep='all'}

LRR_joint_table <- bind_rows(
  Denmark_LRR_result,
  Ireland_LRR_result,
  Norway_LRR_result,
  Sweden_B_LRR_result,
  Sweden_LH_LRR_result
)

pH <- Magge_joint_16S_metadata_filtered_gathered %>%
  filter(Parameter == "Soil_pH") %>%
  select(Treatment, Country, Parameter, Measurement)

pH_normalized <- left_join(LRR_joint_table, pH, by = "Treatment")

# Determing pH scales
pH_normalized <- pH_normalized %>%
  mutate(pH_scale = ifelse(Measurement >= 4 & Measurement <= 5, "4-5",
                    ifelse(Measurement >= 5 & Measurement <= 6, "5-6",
                    ifelse(Measurement >= 6 & Measurement <= 7, "6-7",
                    ifelse(Measurement >= 7 & Measurement <= 8, "7-8", NA))))) %>%
  drop_na() %>%
  filter(Est != 0)

LRR_pH_plot <- ggplot(pH_normalized, aes(reorder(Parameter.x, -Est), y = Est, colour = pH_scale)) + 
  geom_point(size = 2) + 
  geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + 
  labs(y = "LRR 95% confident interval", x = "Parameter", color = "pH scale") +
  theme_light() + 
  scale_colour_viridis_d() + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14,colour = 'black'), 
    strip.background = element_rect(colour = "black"),
    legend.position = c(0.8, 0.2)) +
  facet_wrap(.~Country.x)
  #ggtitle("LRR plot based on soil pH after treatment") + 
  
  

LRR_pH_plot



pH_median <- pH %>%
  mutate(across(Measurement, as.numeric)) %>%
  group_by(Treatment, Country) %>%
  summarise_at(vars(Measurement), funs(median)) %>%
  arrange(Country)

```

```{r LRR functional single plots, Denmark}
# Filter countries

Magge_joint_16S_metadata_filtered_func <- Magge_joint_16S_metadata %>%
  select(X.SampleID, Country, Soil_pH, Treatment, bac:coma) %>%
  filter(Country %in% c("Denmark", "Sweden_LH", "Sweden_B", "Norway", "Ireland"))

# Gathering
Magge_joint_16S_metadata_filtered_gathered_func <- Magge_joint_16S_metadata_filtered_func %>%
  relocate(Treatment, .before = Soil_pH) %>%
  gather(Parameter, Measurement, Soil_pH:coma, factor_key = TRUE)


# Writing functions for LRR
LRR_MAGGE <- function(x = Magge_joint_16S_metadata_filtered_gathered_func, country, treatment, parameter) {

Controls <- Magge_joint_16S_metadata_filtered_gathered_func %>%
  filter(Country == country) %>%
  select(Measurement, Treatment, Parameter) %>%
  filter(Treatment == "Control" & Parameter == parameter) %>%
  select(Measurement)

Response <- Magge_joint_16S_metadata_filtered_gathered_func %>%
  filter(Country == country) %>%
  select(Measurement, Treatment, Parameter) %>%
  filter(Treatment == treatment & Parameter == parameter) %>%
  select(Measurement)

LRRd(as.numeric(Controls$Measurement), as.numeric(Response$Measurement))

}

Denmark_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered_func, y) {
  bind_rows(LRR_MAGGE(x, "Denmark", "4 t/ha lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Denmark", "8 t/ha lime", y),
            LRR_MAGGE(x, "Denmark", "12 t/ha lime", y))
}

# Making a table with results of LRR
Denmark_LRR_result <- bind_rows(
          Denmark_LRR(y = "bac"),
          Denmark_LRR(y = "cren"),
          Denmark_LRR(y = "nirk"),
          Denmark_LRR(y = "nirs"),
          Denmark_LRR(y = "noszi"),
          Denmark_LRR(y = "noszii"),
          Denmark_LRR(y = "aoa"),
          Denmark_LRR(y = "aob"),
          Denmark_LRR(y = "coma")
          )

# Adding columns
Denmark_LRR_result$Parameter <- as.factor(rep(c("bac", "cren", "nirk", "nirs", "noszi", "noszii", "aoa", "aob", "coma"), each = 3))

Denmark_LRR_result$Treatment <- as.factor(rep(c("4 t/ha lime", "8 t/ha lime", "12 t/ha lime")))
Denmark_LRR_result$Treatment <- factor(Denmark_LRR_result$Treatment, levels = c("4 t/ha lime", "8 t/ha lime", "12 t/ha lime"))
Denmark_LRR_result$Country <- rep(c("Denmark"))

# Removing not significant values
Denmark_LRR_result_func <- Denmark_LRR_result %>%
  filter(Est != 0)



# LRR plot                      
Denmark_LRR_result_plot_func <- ggplot(Denmark_LRR_result_func, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_color_viridis_d() + # scale_color_wa_d("skagit") + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Denmark")
  

Denmark_LRR_result_plot_func
```

```{r  LRR functional single plots, Ireland}

Ireland_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered_func, y) {
  bind_rows(LRR_MAGGE(x, "Ireland", "Lime (2011)", y), # Change country and treatment here
            LRR_MAGGE(x, "Ireland", "Lime (2011, 2013)", y),
            LRR_MAGGE(x, "Ireland", "Lime (2013)", y))
}

# Making a table with results of LRR
Ireland_LRR_result <- bind_rows(
          Ireland_LRR(y = "bac"),
          Ireland_LRR(y = "cren"),
          Ireland_LRR(y = "nirk"),
          Ireland_LRR(y = "nirs"),
          Ireland_LRR(y = "noszi"),
          Ireland_LRR(y = "noszii"),
          Ireland_LRR(y = "aoa"),
          Ireland_LRR(y = "aob"),
          Ireland_LRR(y = "coma")
          )

# Adding columns
Ireland_LRR_result$Parameter <- as.factor(rep(c("bac", "cren", "nirk", "nirs", "noszi", "noszii", "aoa", "aob", "coma"), each = 3))


Ireland_LRR_result$Treatment <- as.factor(rep(c("Lime (2011)", "Lime (2011, 2013)", "Lime (2013)")))
Ireland_LRR_result$Treatment <- factor(Ireland_LRR_result$Treatment, levels = c("Lime (2011)", "Lime (2011, 2013)", "Lime (2013)"))
Ireland_LRR_result$Country <- rep(c("Ireland"))
Ireland_LRR_result_func <- Ireland_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Ireland_LRR_result_plot_func <- ggplot(Ireland_LRR_result_func, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_color_viridis_d() + # scale_color_wa_d("skagit") +
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Ireland")
  

Ireland_LRR_result_plot_func
```

```{r single LRR plot Norway}

Norway_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Norway", "Dolomite", y), # Change country and treatment here
            LRR_MAGGE(x, "Norway", "Larvikite", y),
            LRR_MAGGE(x, "Norway", "Marble", y),
            LRR_MAGGE(x, "Norway", "Norite", y),
            LRR_MAGGE(x, "Norway", "Olivine", y))
}

# Making a table with results of LRR
Norway_LRR_result <- bind_rows(
          Norway_LRR(y = "bac"),
          Norway_LRR(y = "cren"),
          Norway_LRR(y = "nirk"),
          Norway_LRR(y = "nirs"),
          Norway_LRR(y = "noszi"),
          Norway_LRR(y = "noszii"),
          Norway_LRR(y = "aoa"),
          Norway_LRR(y = "aob"),
          Norway_LRR(y = "coma")
          )

# Adding columns
Norway_LRR_result$Parameter <- as.factor(rep(c("bac", "cren", "nirk", "nirs", "noszi", "noszii", "aoa", "aob", "coma"), each = 5))


Norway_LRR_result$Treatment <- as.factor(rep(c("Dolomite", "Larvikite", "Marble", "Norite", "Olivine")))
Norway_LRR_result$Treatment <- factor(Norway_LRR_result$Treatment, levels = c("Dolomite", "Larvikite", "Marble", "Norite", "Olivine"))
Norway_LRR_result$Country <- rep(c("Norway"))
Norway_LRR_result_func <- Norway_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Norway_LRR_result_plot_func <- ggplot(Norway_LRR_result_func, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_color_viridis_d() + # scale_color_wa_d("skagit") +
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Norway")
  

Norway_LRR_result_plot_func
```

```{r single LRR plots, Sweden B}

Sweden_B_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Sweden_B", "10 ton/ha lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Sweden_B", "20 ton /ha lime", y),
            )
}

# Making a table with results of LRR
Sweden_B_LRR_result <- bind_rows(
          Sweden_B_LRR(y = "bac"),
          Sweden_B_LRR(y = "cren"),
          Sweden_B_LRR(y = "nirk"),
          Sweden_B_LRR(y = "nirs"),
          Sweden_B_LRR(y = "noszi"),
          Sweden_B_LRR(y = "noszii"),
          Sweden_B_LRR(y = "aoa"),
          Sweden_B_LRR(y = "aob"),
          Sweden_B_LRR(y = "coma")
          )

# Adding columns
Sweden_B_LRR_result$Parameter <- as.factor(rep(c("bac", "cren", "nirk", "nirs", "noszi", "noszii", "aoa", "aob", "coma"), each = 2))


Sweden_B_LRR_result$Treatment <- as.factor(rep(c("10 ton/ha lime", "20 ton/ha lime")))
Sweden_B_LRR_result$Treatment <- factor(Sweden_B_LRR_result$Treatment, levels = c("10 ton/ha lime", "20 ton/ha lime"))
Sweden_B_LRR_result$Country <- rep(c("Sweden_B"))
Sweden_B_LRR_result_func <- Sweden_B_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Sweden_B_LRR_result_plot_func <- ggplot(Sweden_B_LRR_result_func, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_color_viridis_d() + # scale_color_wa_d("skagit") +
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Sweden_B")
  

Sweden_B_LRR_result_plot_func
```

```{r single LRR plots, Sweden LH}

Sweden_LH_LRR <- function(x = Magge_joint_16S_metadata_filtered_gathered, y) {
  bind_rows(LRR_MAGGE(x, "Sweden_LH", "Mixed lime", y), # Change country and treatment here
            LRR_MAGGE(x, "Sweden_LH", "Slaked lime", y),
            LRR_MAGGE(x, "Sweden_LH", "Tunnel kiln slag", y)
            )
}

# Making a table with results of LRR
Sweden_LH_LRR_result <- bind_rows(
          Sweden_LH_LRR(y = "bac"),
          Sweden_LH_LRR(y = "cren"),
          Sweden_LH_LRR(y = "nirk"),
          Sweden_LH_LRR(y = "nirs"),
          Sweden_LH_LRR(y = "noszi"),
          Sweden_LH_LRR(y = "noszii"),
          Sweden_LH_LRR(y = "aoa"),
          Sweden_LH_LRR(y = "aob"),
          Sweden_LH_LRR(y = "coma")
          )

# Adding columns
Sweden_LH_LRR_result$Parameter <- as.factor(rep(c("bac", "cren", "nirk", "nirs", "noszi", "noszii", "aoa", "aob", "coma"), each = 3))


Sweden_LH_LRR_result$Treatment <- as.factor(rep(c("Mixed lime", "Slaked lime", "Tunnel kiln slag")))
Sweden_LH_LRR_result$Treatment <- factor(Sweden_LH_LRR_result$Treatment, levels = c("Mixed lime", "Slaked lime", "Tunnel kiln slag"))
Sweden_LH_LRR_result$Country <- rep(c("Sweden_LH"))
Sweden_LH_LRR_result_func <- Sweden_LH_LRR_result %>%
  filter(Est != 0)


# LRR plot                      
Sweden_LH_LRR_result_plot_func <- ggplot(Sweden_LH_LRR_result_func, aes(reorder(Parameter, -Est), y = Est, colour = Treatment)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + ylab("LRR 95% confident interval") + xlab("Parameter") +
  theme_light() + scale_color_viridis_d() + # scale_color_wa_d("skagit") + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  ggtitle("Sweden_LH")
  

Sweden_LH_LRR_result_plot_func
```

```{r treatments as pH, echo=FALSE,results='hide',fig.keep='all'}

LRR_joint_table_func <- bind_rows(
  Denmark_LRR_result_func,
  Ireland_LRR_result_func,
  Norway_LRR_result_func,
  Sweden_B_LRR_result_func,
  Sweden_LH_LRR_result_func
)

pH <- Magge_joint_16S_metadata_filtered_gathered %>%
  filter(Parameter == "Soil_pH") %>%
  select(Treatment, Country, Parameter, Measurement)

pH_normalized <- left_join(LRR_joint_table_func, pH, by = "Treatment")

# Determing pH scales
pH_normalized <- pH_normalized %>%
  mutate(pH_scale = ifelse(Measurement >= 4 & Measurement <= 5, "4-5",
                    ifelse(Measurement >= 5 & Measurement <= 6, "5-6",
                    ifelse(Measurement >= 6 & Measurement <= 7, "6-7",
                    ifelse(Measurement >= 7 & Measurement <= 8, "7-8", NA))))) %>%
  drop_na() %>%
  filter(Est != 0)

LRR_func_pH_plot_func <- ggplot(pH_normalized, aes(reorder(Parameter.x, -Est), y = Est, colour = pH_scale)) + 
  geom_point(size = 2) + geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) + 
  coord_flip() + labs(y = "LRR 95% confident interval", x = "Parameter", color = "pH scale") +
  theme_light() + scale_color_viridis_d() + #+ scale_color_wa_d("skagit")
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme(
    axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=14, angle = 45), 
    axis.text.y = element_text(colour = "black", size = 14), 
    axis.title.y = element_text(size = 14), 
    strip.text.x = element_text(face = "bold", size = 14,colour = 'black'), 
    strip.background = element_rect(colour = "black"),
    legend.position = c(0.8, 0.2)) +
  facet_wrap(.~Country.x)
  #ggtitle("LRR plot for functional data based on soil pH after treatment") 

LRR_func_pH_plot_func

```

