---
title: "Joint paper MAGGE-pH ITS"
author: ""
date: ""
output: html_document
---
  
Data analysis for joint paper combining functional, ITS and chemical data. 

```{r global_options, echo=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.align = 'center', results = 'hide', echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)
```


```{r load data}

library("phyloseq")

Magge_joint_ITS <- readRDS("path/MAGGE_new_ITS_physeq.rds")

sample_data(Magge_joint_ITS)$Average_yearly_rainfall = gsub("na", "531.93", sample_data(Magge_joint_ITS)$Average_yearly_rainfall)

sample_data(Magge_joint_ITS)$Average_yearly_rainfall <- as.numeric(sample_data(Magge_joint_ITS)$Average_yearly_rainfall)


library(dplyr)
library(stringr)
library(microViz)


# Removing numbers from strings
replacement_string <- c("e1" = "e", "e2" = "e", "e3" = "e", "e4" = "e",
                        "l1" = "l", "l2" = "l", "l3" = "l", "l4" = "l",
                        "finely ground limestone" = "lime", "No lime" = "Control",
                        "0 t/ha lime" = "Control")

Magge_joint_ITS <- Magge_joint_ITS %>%
  ps_mutate(Treatment = str_replace_all(Treatment, replacement_string))

#Remove samples with missing data
Magge_joint_ITS<-ps_drop_incomplete(Magge_joint_ITS, vars = 'Soil_pH', verbose = FALSE)
Magge_joint_ITS_metadata<-data.frame(sample_data(Magge_joint_ITS))

Magge_joint_ITS_metadata <- Magge_joint_16S_metadata %>%
  mutate(Country_code = Country) %>%
  mutate(Country_code = str_replace(Country_code, "France_LQ", "FR1")) %>%
  mutate(Country_code = str_replace(Country_code, "France_TX", "FR2")) %>%
  mutate(Country_code = str_replace(Country_code, "Denmark", "DK")) %>%
  mutate(Country_code = str_replace(Country_code, "Finland", "FI")) %>%
  mutate(Country_code = str_replace(Country_code, "Ireland", "IE")) %>%
  mutate(Country_code = str_replace(Country_code, "New_Zealand", "NZ")) %>%
  mutate(Country_code = str_replace(Country_code, "Norway", "NO")) %>%
  mutate(Country_code = str_replace(Country_code, "Sweden_B", "SE1")) %>%
  mutate(Country_code = str_replace(Country_code, "Sweden_LH", "SE2"))


```

Rarefaction curves to  determine subsampling level. 28000 reads chosen. All samples used for analysis. Data was then subsampled to only use control sites (no pH treatment/liming)

```{r rarefaction}

# Remove any samples with less than 100 sequences
pruned <- prune_samples(sample_sums(Magge_joint_ITS) >= 100, Magge_joint_ITS)
pruned

library(ggplot2)
library(ranacapa)

# Calculate rarefaction curves, then plot and color by subsite
p <- ggrare(pruned, step = 1000, color = "Country", se =
              FALSE,parallel = TRUE)+
  scale_y_log10() +
  scale_x_continuous(limits = c(1, 170000))
p


p + facet_wrap(~Country)

# Rarefy at 28k

Magge_joint_ITS_rare <- rarefy_even_depth(Magge_joint_ITS, 28000, rngseed=TRUE)
```

NMDS ordinations based on ITS data
ANOSIM and ADONIS results shown below done by Country and then pH, with an additional Mantel test for pH at the end.
```{r Compare beta diversity, results='markup'}

# Ordinate data using Non-metric multidimensional scaling (NMDS) on Bray–Curtis dissimilarity (distances)
set.seed(1)
NMDS_ITS.ord <- ordinate(Magge_joint_ITS_rare, "NMDS", "bray")

# By calling the newly created file we can get the stress value of the plot
NMDS_ITS.ord

library(vegan)
library(grid)

stressplot(NMDS_ITS.ord)

# Create label for stress value
Stress_val = grobTree(textGrob("Stress = 0.17", x = 0.05,  y = 0.95, hjust = 0, gp = gpar(col = "Black", fontsize = 14,fontface = "italic")))

library(ggrepel)
library(wacolors)

options(ggrepel.max.overlaps = Inf)


Magge_ITS_NMDS <- plot_ordination(Magge_joint_ITS_rare, NMDS_ITS.ord, type = "samples",color="Soil_pH")+
  geom_text_repel(aes(label = Magge_joint_ITS_metadata$Country_code)) + 
  scale_color_wa_c(wacolors$vantage, reverse=TRUE)+
  geom_point(size = 1) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(color = "Soil pH")+
  annotation_custom(Stress_val)+
  theme_bw(base_size=12)



Magge_ITS_NMDS



# Stats by Country
# ANOSIM
Country_group = get_variable(Magge_joint_ITS_rare, "Country")
Country_ano = anosim(phyloseq::distance(Magge_joint_ITS_rare, "bray"), Country_group)
Country_ano$signif
Country_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_ITS_rare), "data.frame")
# Calculate your Bray distance matrix
Country_ado = phyloseq::distance(Magge_joint_ITS_rare, "bray")
# Perform your ADONIS test
Country_ado_stats <- adonis2(Country_ado ~ Country, df_ado)

# Check results
Country_ado_stats


# Stats by pH
# ANOSIM
pH_group = get_variable(Magge_joint_ITS_rare, "Soil_pH")
pH_ano = anosim(phyloseq::distance(Magge_joint_ITS_rare, "bray"), pH_group)
pH_ano$signif
pH_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_ITS_rare), "data.frame")
# Calculate your Bray distance matrix
pH_ado = phyloseq::distance(Magge_joint_ITS_rare, "bray")
# Perform your ADONIS test
pH_ado_stats <- adonis2(pH_ado ~ Soil_pH, df_ado)

# Check results
pH_ado_stats

# Mantel correlation
### pH matrix
ph.dist <- dist(df_ado$Soil_pH, method = "euclidean")
dist.matrix <- as.matrix(ph.dist)

### Mantel tests
mantel(pH_ado, ph.dist, method = "spearman")

```


NMDS color based on rain

```{r NMDS color based on rain, results='markup'}

Magge_ITS_NMDS_rain <- plot_ordination(Magge_joint_ITS_rare, NMDS_ITS.ord, type = "samples",color="Average_yearly_rainfall")+
  geom_text_repel(aes(label = Magge_joint_ITS_metadata$Country_code)) + 
  scale_color_wa_c(wacolors$vantage, reverse=TRUE)+
  geom_point(size = 1) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(colour = "black", size=12), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(color = "Average yearly rainfall")+
  annotation_custom(Stress_val)+
  theme_bw(base_size=12)

Magge_ITS_NMDS_rain


# Stats by rain
# ANOSIM
rain_group = get_variable(Magge_joint_ITS_rare, "Average_yearly_rainfall")
rain_ano = anosim(phyloseq::distance(Magge_joint_ITS_rare, "bray"), rain_group)
rain_ano$signif
rain_ano$statistic

# ADONIS
# Create a data frame using your sample_data
df_ado = as(sample_data(Magge_joint_ITS_rare), "data.frame")
# Calculate your Bray distance matrix
rain_ado = phyloseq::distance(Magge_joint_ITS_rare, "bray")
# Perform your ADONIS test
rain_ado_stats <- adonis2(rain_ado ~ Average_yearly_rainfall, df_ado)

# Check results
rain_ado_stats

# Mantel correlation
### Rain matrix
rain.dist <- dist(df_ado$Average_yearly_rainfall, method = "euclidean")
dist.matrix <- as.matrix(rain.dist)

### Mantel tests
mantel(rain_ado, rain.dist, method = "spearman")
```


Stacked bar plot based on phylum (draft and clean version)

```{r stack bar phylum}

library(plyr)
library(scales)

# Clean graph and cluster by country

library(dplyr)
library(microViz)
library(fantaxtic)
library(speedyseq)
library(phylosmith)

Magge_joint_ITS_rare <- merge_treatments(Magge_joint_ITS_rare, c("Country", "Treatment"))

MJITSR_clean <- name_na_taxa(Magge_joint_ITS_rare, na_label = "Unidentified <tax> (<rank>)")

MJITSR_clean2 <- MJITSR_clean %>%
  ps_select(Country, Treatment, Country_Treatment, Soil_pH) %>% # avoids lots of phyloseq::merge_samples warnings
  merge_samples2(group = "Country_Treatment", funs = list()) %>%
  tax_glom(taxrank = "Phylum")

MJITSR_clean2 %>%
  tax_filter(
    tax_level = "Phylum", min_prevalence = 0.001,
    prev_detection_threshold = 100
  ) %>%
  ps_mutate(
    Treatment = factor(Treatment, levels = c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)','Dolomite', 'Larvikite', 'Marble','Norite','Olivine','10 ton/ha lime', '20 ton /ha lime','Slaked lime', 'Tunnel kiln slag', 'Mixed lime'))) %>%  
  ps_mutate(
    Country = factor(Country, levels = c('Denmark', 'Ireland', 'Finland', 'Norway', 'New_Zealand', 'France_LQ', 'Sweden_B', 'France_TX', 'Sweden_LH'))) %>%  
  comp_barplot(
    tax_level = "Phylum", n_taxa = 10,
    bar_outline_colour = NA,
    bar_width = 0.7,
    taxon_renamer = toupper,
    facet_by = "Country",
    label = "Treatment",
    x = "Treatment") + 
  theme(legend.text = element_text(size = 12),
        legend.title =element_text(size = 14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(angle = 45, colour = "black", size = 10, vjust = 1, hjust = 1), 
        axis.text.y = element_text(colour = "black", size = 12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())



```

Stacked bar plot based on Order (clean version)

```{r stack bar Order}

# Clean graph and cluster by country

MJITSR_clean2_Order<-MJITSR_clean%>%
  ps_select(Country, Treatment, Country_Treatment, Soil_pH) %>% # avoids lots of phyloseq::merge_samples warnings
  merge_samples2(group = "Country_Treatment", funs = list())%>%
  tax_glom(taxrank = "Order")

MJITSR_clean2_Order %>%
  tax_filter(
    tax_level = "Order", min_prevalence = 0.001,
    prev_detection_threshold = 100
  ) %>%
  ps_mutate(
    Treatment = factor(Treatment, levels = c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)','Dolomite', 'Larvikite', 'Marble','Norite','Olivine','10 ton/ha lime', '20 ton /ha lime','Slaked lime', 'Tunnel kiln slag', 'Mixed lime'))) %>%  
  ps_mutate(
    Country = factor(Country, levels = c('Denmark', 'Ireland', 'Finland', 'Norway', 'New_Zealand', 'France_LQ', 'Sweden_B', 'France_TX', 'Sweden_LH'))) %>%  
  comp_barplot(
    tax_level = "Order", n_taxa = 15,
    bar_outline_colour = NA,
    bar_width = 0.7,
    taxon_renamer = toupper,
    facet_by = "Country",
    label = "Treatment",
    x = "Treatment"
  ) + 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size =14, face="bold"),
        axis.title = element_text(size = 14), 
        axis.text.x = element_text(angle=45, colour = "black", size=10, vjust = 1, hjust = 1), 
        axis.text.y = element_text(colour = "black", size=12),
        panel.background = element_rect(fill = "grey98"),
        panel.grid = element_line(color = "grey75"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank())

```



distance comparisons Ireland

```{r distance comparisons Ireland}

Magge_joint_ITS_rare_IR <- subset_samples(Magge_joint_ITS_rare, Country == "Ireland")
Magge_joint_ITS_rare_IR

Totu_table = t(otu_table(Magge_joint_ITS_rare_IR)) # transpose otu table
otu_table(Magge_joint_ITS_rare_IR) = Totu_table

p = Magge_joint_ITS_rare_IR
m = "bray"
s = "X.SampleID"
d = "Treatment"   # Day

# Calculate distances
library(reshape2)
library(reshape)
library(dplyr)


wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor, as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor, as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_its_ireland = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Ireland")

library(ggpubr)

# Order factors
wu.sd.sums_its_ireland$Treatment2 = factor(wu.sd.sums_its_ireland$Treatment2, levels = c('Control','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)'))

IR_ITS_dist_plot<-ggplot(wu.sd.sums_its_ireland, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") +
  theme(axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size = 18), axis.title.y = element_text(size = 18), 
        strip.text.x = element_text(face = "bold", size = 18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size = 3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control") + ylim(0, 1)+ggtitle("Ireland")

IR_ITS_dist_plot


```

distance comparisons Norway
```{r distance comparisons Norway}

Magge_joint_ITS_rare_NO <- subset_samples(Magge_joint_ITS_rare, Country == "Norway")
Magge_joint_ITS_rare_NO

Totu_table =t(otu_table(Magge_joint_ITS_rare_NO)) #transpose otu table
otu_table(Magge_joint_ITS_rare_NO)=Totu_table

p = Magge_joint_ITS_rare_NO
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calc distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_its_norway = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Norway")


# Order factors
wu.sd.sums_its_norway$Treatment2 = factor(wu.sd.sums_its_norway$Treatment2, levels=c('Control','Dolomite', 'Larvikite', 'Marble','Norite','Olivine'))

NO_ITS_dist_plot<-ggplot(wu.sd.sums_its_norway, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Norway")
NO_ITS_dist_plot


```

distance comparisons Denmark

```{r distance comparisons Denmark}

Magge_joint_ITS_rare_DE <- subset_samples(Magge_joint_ITS_rare, Country == "Denmark")
Magge_joint_ITS_rare_DE

Totu_table =t(otu_table(Magge_joint_ITS_rare_DE)) #transpose otu table
otu_table(Magge_joint_ITS_rare_DE)=Totu_table

p = Magge_joint_ITS_rare_DE
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculating distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Removing self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_its_denmark = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Denmark")


# Order factors
wu.sd.sums_its_denmark$Treatment2 = factor(wu.sd.sums_its_denmark$Treatment2, levels=c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))


DE_ITS_dist_plot<-ggplot(wu.sd.sums_its_denmark, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30, aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Denmark")

DE_ITS_dist_plot

```


distance comparisons Sweden_LH

```{r distance comparisons Sweden_LH}

Magge_joint_ITS_rare_SWLH <- subset_samples(Magge_joint_ITS_rare, Country == "Sweden_LH")
Magge_joint_ITS_rare_SWLH

Totu_table =t(otu_table(Magge_joint_ITS_rare_SWLH)) #transpose otu table
otu_table(Magge_joint_ITS_rare_SWLH)=Totu_table

p = Magge_joint_ITS_rare_SWLH
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculate distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_its_swedenlh = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Sweden_LH")




SWLH_ITS_dist_plot<-ggplot(wu.sd.sums_its_swedenlh, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x = element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30,  aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Sweden_LH")

SWLH_ITS_dist_plot

```


distance comparisons Sweden_B

```{r distance comparisons Sweden_B}
Magge_joint_ITS_rare_SWB <- subset_samples(Magge_joint_ITS_rare, Country == "Sweden_B")
Magge_joint_ITS_rare_SWB

Totu_table =t(otu_table(Magge_joint_ITS_rare_SWB)) #transpose otu table
otu_table(Magge_joint_ITS_rare_SWB)=Totu_table

p = Magge_joint_ITS_rare_SWB
m = "bray"
s = "X.SampleID"
d = "Treatment"   

# Calculate distances

wu = vegan::vegdist(t(otu_table(p)), method = "bray")
wu.m = melt(as.matrix(wu))

colnames(wu.m) <- c("Var1", "Var2", "value")

# Remove self-comparisons
wu.m = wu.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)

# Get sample data (S4 error OK and expected)
sd = data.frame(sample_data(p))%>%
  select(s, d)%>%
  mutate_if(is.factor,as.character) 

# Combined distances with sample data
colnames(sd) = c("Var1", "Treatment")
wu.sd = left_join(wu.m, sd, by = "Var1")

colnames(sd) = c("Var2", "Treatment2")
wu.sd = left_join(wu.sd, sd, by = "Var2")

wu.sd.sums_its_swedenb = wu.sd %>%
  filter(Treatment == 'Control')%>%
  mutate_if(is.factor,as.character) %>%
  mutate(country = "Sweden_B")


# Order factors
wu.sd.sums_its_swedenb$Treatment2 = factor(wu.sd.sums_its_swedenb$Treatment2, levels=c('Control','10 ton/ha lime', '20 ton /ha lime'))

SWB_ITS_dist_plot<-ggplot(wu.sd.sums_its_swedenb, aes(x = Treatment2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot() + ylab("Bray distance") + xlab("Treatment") +
  theme(axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=18, angle = 45), axis.text.y = element_text(colour = "black", size=18), axis.title.y = element_text(size=18), 
        strip.text.x = element_text(face="bold", size=18), 
        strip.background = element_rect(colour = "black", fill = "white")) +
  stat_compare_means(method = "kruskal.test", size=3,vjust = 30,  aes(group = Treatment2))+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ylim(0, 1)+ggtitle("Sweden_B")

SWB_ITS_dist_plot

```


```{r Find significantly affected Phyla using rstatix}

#no hits after p adjusted so used just p value
Magge_joint_ITS_rare_DE_df<-Magge_joint_ITS_rare_DE %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 
#how does this community look?
d<- ggplot(Magge_joint_ITS_rare_DE_df,aes(x=Treatment,y=Abundance,fill=Phylum))+ 
  geom_bar(stat="identity",position="fill") +
  scale_y_continuous()+
  ylab("Relative Abundance")
d

library(rstatix)
# do tests
KW_taxa_ITS_DE<-Magge_joint_ITS_rare_DE_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Denmark")



# make a list of the significant genera
only_sig_taxa_ITS_DE<-Magge_joint_ITS_rare_DE_df %>% 
  filter(OTU %in% KW_taxa_ITS_DE$OTU) 


only_sig_taxa_ITS_DE$Treatment = factor(only_sig_taxa_ITS_DE$Treatment, levels=c('Control','4 t/ha lime', '8 t/ha lime', '12 t/ha lime'))

# plot
DE_sig_tax_plot_ITS<-ggplot(only_sig_taxa_ITS_DE, aes(x=Treatment, y=Abundance, fill=Phylum)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ggtitle("Denmark")

DE_sig_tax_plot_ITS




#Ireland
#no hits after p adjusted so used just p value
Magge_joint_ITS_rare_IR_df<-Magge_joint_ITS_rare_IR %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_ITS_IR<-Magge_joint_ITS_rare_IR_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Ireland")



# make a list of the significant genera
only_sig_taxa_ITS_IR<-Magge_joint_ITS_rare_IR_df %>% 
  filter(OTU %in% KW_taxa_ITS_IR$OTU) 


only_sig_taxa_ITS_IR$Treatment = factor(only_sig_taxa_ITS_IR$Treatment, levels = c('Control','Lime (2011)', 'Lime (2011, 2013)', 'Lime (2013)'))

# plot
IR_sig_tax_plot_ITS<-ggplot(only_sig_taxa_ITS_IR, aes(x=Treatment, y=Abundance, fill=Phylum)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = "Control")+ggtitle("Ireland")

IR_sig_tax_plot_ITS




#Sweden LH
#no hits after p adjusted so used just p value
Magge_joint_ITS_rare_SWLH_df<-Magge_joint_ITS_rare_SWLH %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_ITS_SWLH<-Magge_joint_ITS_rare_SWLH_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Sweden_LH")



# make a list of the significant genera
only_sig_taxa_ITS_SWLH<-Magge_joint_ITS_rare_SWLH_df %>% 
  filter(OTU %in% KW_taxa_ITS_SWLH$OTU) 


only_sig_taxa_ITS_SWLH$Treatment = factor(only_sig_taxa_ITS_SWLH$Treatment, levels=c('Control','Mixed lime', 'Slaked lime', 'Tunnel kiln slag'))

# plot
SWLH_sig_tax_plot_ITS<-ggplot(only_sig_taxa_ITS_SWLH, aes(x=Treatment, y=Abundance, fill=Phylum, group=Treatment)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = 'Control')+ggtitle("Sweden LH")

SWLH_sig_tax_plot_ITS



#no hits after p adjusted so used just p value
Magge_joint_ITS_rare_SWB_df<-Magge_joint_ITS_rare_SWB %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_ITS_SWB<-Magge_joint_ITS_rare_SWB_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Sweden_B")



# make a list of the significant genera
only_sig_taxa_ITS_SWB<-Magge_joint_ITS_rare_SWB_df %>% 
  filter(OTU %in% KW_taxa_ITS_SWB$OTU) 


only_sig_taxa_ITS_SWB$Treatment = factor(only_sig_taxa_ITS_SWB$Treatment, levels=c('Control','10 ton/ha lime', '20 ton /ha lime'))

# plot
SWB_sig_tax_plot_ITS<-ggplot(only_sig_taxa_ITS_SWB, aes(x=Treatment, y=Abundance, fill=Phylum, group=Treatment)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = 'Control')+ggtitle("Sweden LH")

SWB_sig_tax_plot_ITS



#no hits after p adjusted so used just p value
Magge_joint_ITS_rare_NO_df<-Magge_joint_ITS_rare_NO %>% 
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
  psmelt() 


# do tests
KW_taxa_ITS_NO<-Magge_joint_ITS_rare_NO_df%>% 
  group_by(OTU) %>% 
  kruskal_test(Abundance ~ Treatment) %>% 
  adjust_pvalue(method="fdr")%>% 
  filter(p < 0.05) %>%
  mutate(Country = "Norway")



# make a list of the significant genera
only_sig_taxa_ITS_NO<-Magge_joint_ITS_rare_NO_df %>% 
  filter(OTU %in% KW_taxa_ITS_NO$OTU) 


only_sig_taxa_ITS_NO$Treatment = factor(only_sig_taxa_ITS_NO$Treatment, levels=c('Control','Dolomite', 'Larvikite', 'Marble','Norite','Olivine'))

# plot
NO_sig_tax_plot_ITS<-ggplot(only_sig_taxa_ITS_NO, aes(x=Treatment, y=Abundance, fill=Phylum, group=Treatment)) + 
  geom_boxplot() + 
  facet_wrap(Phylum~Class,scales="free_y")+
  theme(
    axis.text.x=element_text(colour = "black", vjust = 1, hjust = 1, size=12, angle = 45), 
    axis.text.y = element_text(colour = "black", size=12), 
    axis.title.y = element_text(size=12),
    strip.text.x = element_text(size=10), 
    strip.background = element_rect(colour = "black", fill = "white"))+
  stat_compare_means(method = "kruskal.test", paired = FALSE)+
  stat_compare_means(aes(label = after_stat(p.signif)),
                     method = "t.test", ref.group = 'Control')+ggtitle("Norway")

NO_sig_tax_plot_ITS




# Combined table for taxa
KW_taxa_ITS_all <- bind_rows(KW_taxa_ITS_DE, KW_taxa_ITS_IR, KW_taxa_ITS_SWB, KW_taxa_ITS_SWLH,KW_taxa_ITS_NO)
KW_taxa_ITS_all <- KW_taxa_ITS_all %>%
  select(-.y., -n, -df) %>%
  arrange(desc(p))

KW_taxa_ITS_all

```

